\texorpdfstring {Implementing Equational Constraints \\ in a Functional Language} {Implementing Equational Constraints in a Functional Language}

### 2012-09-24 papers/wlp2011/paper.tex
Sentence about practical usage added

We have presented an implementation of equational constraints
and functional patterns in KiCS2, a purely functional implementation
of Curry.
In addition to the kernel implementation described in
\cite{BrasselHanusPeemoellerReck11},
we add binding constraints
to computed values which are processed when values are extracted
at the top level of a computation.
Since only new constructors and pattern matching rules 
for them are added in our implementation,
no overhead is introduced for programs
without equational constraints, i.e.,
our implementation does not sacrifice the high efficiency
of the kernel implementation shown in
\cite{BrasselHanusPeemoellerReck11}.
However, if these features are used, they usually lead
to a comparably efficient execution, as demonstrated by our benchmarks.

Other implementations of equational constraints in functional logic
languages use side effects for their implementation.
For instance, PAKCS \cite{Hanus10PAKCS}
exploits the implementation of logic variables in Prolog,
which are implemented on the primitive level by side effects.
MCC \cite{Lux99FLOPS} compiles into C where a specific
abstract machine implements the handling of logic variables.
We have shown that our implementation is competitive to those.
In contrast to those systems, our implementation supports
a variety of ``top-level'' search strategies,
like iterative deepening, breadth-first or parallel search,
as well as user-programmable search strategies,
where the avoidance of side effects is important.

>>>>

We have presented an implementation of equational constraints
and functional patterns in KiCS2, a purely functional implementation
of Curry.
In addition to the kernel implementation described in
\cite{BrasselHanusPeemoellerReck11},
we add binding constraints
to computed values which are processed when values are extracted
at the top level of a computation.
Since only new constructors and pattern matching rules 
for them are added in our implementation,
no overhead is introduced for programs
without equational constraints, i.e.,
our implementation does not sacrifice the high efficiency
of the kernel implementation shown in
\cite{BrasselHanusPeemoellerReck11}.
However, if these features are used, they usually lead
to a comparably efficient execution, as demonstrated by our benchmarks.
Although the benchmarks were small in order to evaluate
our unification implementation,
it should be noted that KiCS2 is used in larger applications,
like the curricula and module information system of our department%
\footnote{\url{http://www-ps.informatik.uni-kiel.de/~mh/studiengaenge/}}.
In these and similar applications, where large parts are
purely functional computations,
KiCS2 is 15-20 times faster than PAKCS \cite{Hanus10PAKCS}.

Other implementations of equational constraints in functional logic
languages are based on side effects.
For instance, PAKCS \cite{Hanus10PAKCS}
exploits the implementation of logic variables in Prolog,
which are implemented on the primitive level by side effects.
MCC \cite{Lux99FLOPS} compiles into C where a specific
abstract machine implements the handling of logic variables.
We have shown that our implementation is competitive to those.
In contrast to those systems, our implementation supports
a variety of ``top-level'' search strategies,
like iterative deepening, breadth-first or parallel search,
as well as user-programmable search strategies,
where the avoidance of side effects is important.
